\documentclass{article}
\usepackage{standardStyle}
\begin{document}
Lets see how this works.



\section{2017/7/3, Yang}
\label{sec:201773-yang}

I'm currently working on heuristics and related papers. Below are some of my thoughts on your comments. Let me do it one by one:

\begin{enumerate}
\item As for the learner, I think you're right. I also thought of this idea as we've actually discussed it last time. However, I don't really get how to proceed the "guiding" or the "teaching" part. To me, this is something like letting the learner adapt the probability weights on different heuristics. Perhaps I should read more papers on it. BTW thanks for the Q-learning material, this is exactly the topic I've been interested in (but a little bit wavering whether to step into).

7/4/N: I'll put the q learning files here too. Then you copy what you need into the ward.tex file, and throw away what has no use anymore.

I think that with the heuristics it must be possible to set the initial probabilities. See my updates in the eoq.tex file, last section. To learn this stuff, we might first work on inventory models. That is easy at least.

\item The story. Yeah if I'm not wrong then we've come up with a nice story behind this game. What I mean was to properly elaborate it, or generalize it. Perhaps the game is not only for this kind of scheduling problem. I'm not certain on it.


7/4/N: For the story, we need to explain the real problem. I'll do this. 

\item One last thing, but does matter: as I said, the connection through my VPN to out of Chinese GFW is unstable, and thus I'm afraid I cannot react upon your mails instantly every time. 


7/4/N. No problem. Perhaps working like this with github works also fine. Lets see how it goes.

For this reason, I'm wondering whether you'd like to download an app called "WeChat", which is like the Whatsapp but proved to work fine across the GFW. It's a Chinese app but is on all app stores globally, so you can simply search and download it freely. If so, then we can contact each other instantly and handle most small chats through WeChat.

7/4/N: does it work on linux too? 

\end{enumerate}

\section{Action points}
\label{sec:action-points}

\begin{enumerate}
\item N: check student thesis of Durk-Jouke on ward scheduling to provide background/intro for our work.
\end{enumerate}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
